Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm

### 2022-02-10 2022/PAVT/PAVT2022.tex
fix reviewer comment in the UNN part

\documentclass{svproc}
%
% RECOMMENDED 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}

@@

\begin{equation}
  k_i (T) = k_i^0 \exp \left(- \dfrac{E_i}{RT} \right),
  \label{eq:arren}
\end{equation}
where $k_i(T)$ is a constant of $i$-th stage of reation rate, $k_i^0$ is a pre-exponential factor of $i$-th reation stage, $E_i$ is an activation energy, J/mol, $R$ is the universal gas constant, J $\cdot$ (K$\cdot$mol), $T$ is the temperature, K. Thus, in order to fully develop the kinetic model of the reaction, it is necessary to calculate the activation energies $E_i$ and pre-exponential factors $k_i^0$ of all stages of the chemical reaction.  There are two formulations of the problems of searching for kinetic parameter data $E$ and $k^0$. The first one is to solve the inverse problem of selecting the kinetic parameters included in the equation (\ref{eq:arren}), which during the solution allow us to calculate all the reaction rate constants $k_i$. Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one})--(\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data. Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function:

@@

As it has been already mentioned above, from the formal point of view, we consider the inverse problem of chemical kinetics as Lipschitz global optimization problem. In the general form, a problem of the class specified above can be formulated mathematically as 
follows:
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
where $a,b$ are given vectors, $a,b\in R^N$, and the objective function $\varphi(y)$ satisfies the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

@@

@@

According to the problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
%ННГУ
To carry out the numerical experiments, UNN supercomputer Lobachevsky was used (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node). The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.

@@

>>>>

\documentclass{svproc}
%
% RECOMMENDED 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{color}

@@

\begin{equation}
  k_i (T) = k_i^0 \exp \left(- \dfrac{E_i}{RT} \right),
  \label{eq:arren}
\end{equation}
where $k_i(T)$ is a constant of $i$-th stage of reation rate, $k_i^0$ is a pre-exponential factor of $i$-th reation stage, $E_i$ is an activation energy, J/mol, $R$ is the universal gas constant, J $\cdot$ (K$\cdot$mol), $T$ is the temperature, K. Thus, in order to fully develop the kinetic model of the reaction, it is necessary to calculate the activation energies $E_i$ and pre-exponential factors $k_i^0$ of all stages of the chemical reaction.  There are two formulations of the problems of searching for kinetic parameter data $E$ and $k^0$. The first one is to solve the inverse problem of selecting the kinetic parameters included in the equation (\ref{eq:arren}), which during the solution allow us to calculate all the reaction rate constants $k_i$. Then, with the found rate constants of the reaction stages, the system of differential equations (\ref{eq:one})--(\ref{eq:twelve}) is solved, then the calculated concentrations are compared with the corresponding experimental data. Mathematically, this problem has the following formulation: it is necessary to minimize the following objective function

@@

\textcolor[rgb]{1,0,0}{As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
In the specific problem under consideration, the values of the objective function are calculated by solving stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}). Since the right parts of the system are continuous functions with bounded derivatives, then theoretically its solution will also be continuous and bounded. Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.}

In the general form, a problem of the class specified above can be formulated mathematically as 
follows:
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
where $a,b$ are given vectors, $a,b\in R^N$, and the objective function $\varphi(y)$ satisfies the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

@@

%Данный комментарий можно и не включать...
%Note that the presence of rapidly decreasing components in the solution of the stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}) can lead to the fact that in different subdomains the objective function can have different local Lipschitz constants. This problem will affect the convergence of any of the global optimization methods. To take it into account, a number of approaches have been proposed related to local tuning of global optimization methods (see, e.g., \cite{Barkalov2021,Strongin2020}).

@@

According to the problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
%ННГУ
To carry out the numerical experiments, UNN supercomputer Lobachevsky was used (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node). The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
\textcolor[rgb]{1,0,0}{The accuracy of solving the ODE system (\ref{eq:one})--(\ref{eq:twelve}) was set small enough so that the final error of the solution was much less than the accuracy of the stopping criterion of the optimization method and did not affect the method used.}

@@

\textcolor[rgb]{1,0,0}{Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function. The constant is adaptively estimated during the work of the algorithm and may vary depending on the accumulated search information. With a correct estimate of the Lipschitz constant the method can converge to a global optimum point faster than in the case of its incorrect estimate. This explains the effect of superlinear speedup observed in experiments.}
